Global patterns of forest autotrophic carbon fluxes

### 2020-05-05 manuscript/manuscript draft.Rmd
resolved issue #72
closes #72

Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @kunert_understanding_2019]. By constraining our analysis to stands ≥ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.

>>>>

Past studies have differed in their conclusions regarding the relationship between FACF and latitude or its correlates (Table 1, *H1*)--quite possibly because of lack of standardization with respect to stand age and disturbance history. Our findings indicate that, among mature, undisturbed stands, FACF are unambiguously highest in the tropical regions, and the relationship is approximately linear (Fig. 2). This contrasts with the suggestion that productivity of temperate forests is similar to that of tropical forests [@huston_global_2009]. Compared to tropical forests, the temperate forest biome has experienced more widespread anthropogenic disturbance and has a larger fraction of secondary stands [@potapov_mapping_2008; @poulter_global_2018], so analyses comparing across latitudinal gradients without controlling for stand age risk confounding age with biome effects. In addition, because carbon allocation varies with stand age [@delucia_forest_2007] (**Doughty et al. 2018: DOI: 10.1111/btp.12504; Anderson-Teixeira et al. 2013: doi: 10.1111/gcb.12194**), age differences may introduce systematic biases into analyses of FACF across latitude or global climatic gradients. For example, woody productivity tends to be higher in rapidly aggrading secondary stands than in old-growth forests, where proportionally more C is allocated to respiration (**Doughty et al. 2018: DOI: 10.1111/btp.12504**);[@delucia_forest_2007; @kunert_understanding_2019]. By constraining our analysis to stands $\ge$ 100 years old with no record of major recent disturbance, we have clarified the shape of the relationship between FACF and latitude.
